\acknowledgements{%
    \begin{adjustbox}{minipage=.8\textwidth,center}
        \emph{`The best theory is inspired by practice. The best practice is inspired by
        theory.'}

        \hspace*{\fill} --- Donald E. Knuth, \emph{Selected Papers on Computer Science},
        1996

        \bigskip

        \emph{`We have a habit in writing articles published in scientific journals to
        make the work as finished as possible, to cover up all the tracks, to not
        worry about the blind alleys or describe how you had the wrong idea first, and
        so on. So there isn't any place to publish, in a dignified manner, what you
        actually did in order to get to do the work.'}

        \hspace*{\fill} --- Richard P. Feynman, \emph{Nobel Lecture}, 1966

        \bigskip
    \end{adjustbox}

    \bigskip

    First of all, I would like to thank my advisors, Gopal Ramchurn and Danesh Tarapore,
    who allowed me to embark on this path, and helped me to grow both as a researcher and
    as a professional. I am grateful to Klaus-Peter Zauner, Tim Norman and Jie Zhang for
    examining my progress and giving me valuable advice. Thanks also to Alessandro
    Farinelli and Mohammad Divband Soorati for the edifying discussions, and to Ryan Beal
    and Shaun Lamb for the good times we had together in Chania and in the best breweries
    of Southampton.

    This research is supported by the EPSRC grant EP/R2115315/1
    and the \href{https://www.axa-research.org/en/project/sarvapali-gopal-ramchurn}{AXA
    Research Fund}. I would also like to acknowledge the use of the Iridis High
    Performance Computing Facility, and associated support services at the University of
    Southampton.

    To all the people with whom I have made friends, or even shared just a laugh, in
    student societies such as Italian Society, MexSoc, CathSoc, Live Music Society, SLAPS
    and Salsa Society, to name a few. In particular, I would like to mention my carefree
    Sundays of roast, Mexican chocolate and Neapolitan memes with Pierfrancesco and
    Davide, which helped me to feel less homesick and to cope with pressing commitments.

    To the L{\'o}pez de la Cruz family, who welcomed me warmly and allowed me to work in
    peace during uncertain months. To Elizabeth, with whom I went hiking in mountains such
    as Cerro de la Silla, Cerro Agujerado and Cerro de Chipinque, and then ended the day
    by going for chilaquiles. To Ixora and all my other Mexican friends, with whom I
    spent many fun evenings.

    To my family, in particular my mother and my little sister, true forces of nature, who
    have always helped and supported me.

    To my beautiful love Ilse, without whom I would not be here. \emph{Te cielo}.
}

\dedicatory{%
    Dedicated to the women of my life, in order of appearance: mum Agnese,\\
    granny Luigia, aunt Nunzia, little sister Annaclara, and amorcita Ilse.%
}
